Adaptive group-regularized logistic elastic net regression Supplementary material

### 2019-03-06 docs/supplement.Rnw
added briers to rnaseq_lymph_node_metastasis.R and micrornaseq_colorectal_cancer.R

\documentclass[a4paper,hidelinks]{article}
\usepackage{bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,
rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,
bbm,algorithm,algpseudocode,pgfplotstable,threeparttable,booktabs,mathtools,
dsfont,parskip}

@@

% functions and operators
\renewcommand{\L}{\mathcal{L}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\argmax}{\text{argmax} \,}
\newcommand{\expit}{\text{expit}}
\newcommand{\erfc}{\text{erfc}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\tr}{^{\text{T}}}
\newcommand{\diag}{\text{diag}}
\newcommand{\KL}{D_{\text{KL}}}
\newcommand{\trace}{\text{tr}}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\newcommand{\appropto}{\mathrel{\vcenter{
  \offinterlineskip\halign{\hfil$##$\cr
    \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
\newcommand{\card}[1]{\text{card} \left( #1 \right)}
\makeatletter
\newcommand*{\defeq}{\mathrel{\rlap{%
			\raisebox{0.3ex}{$\m@th\cdot$}}%
		\raisebox{-0.3ex}{$\m@th\cdot$}}%
	=}
\makeatother

@@

% settings
\graphicspath{{../figs/}}
\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
\setlength{\textwidth}{15cm}
\setlength{\parindent}{0pt}
\title{Adaptive group-regularized logistic elastic net regression Supplementary 
material}
\date{}
\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: 
\href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Carel F.W. Peeters$^{1}$, 
Aad W. van der Vaart$^{2}$ \\ and Mark A. van de Wiel$^{1,3}$}

@@

  \maketitle
  \noindent
  1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health
  research institute, VU University Medical Center, PO Box 7057, 1007 MB
  Amsterdam, The Netherlands\\
  2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\
  3. Department of Mathematics, VU University, Amsterdam, The Netherlands \\

@@

  \section{Introduction}
  This document contains supplementary material to the paper `Adaptive 
  group-regularized logistic elastic net regression'. Section 
  \ref{sec:enetprior} contains extra details on the elastic net prior 
  distribution. Section \ref{sec:objectivefunction} gives the objective 
  function that is maximised in the variational Bayes procedure introduced in
  Section 3 of the Main Document. Section \ref{sec:polyagamma} gives details on
  the P{\'o}lya-Gamma parametrisation introduced in Section 3 of the Main
  Document. The derivations of the variational posterior used in the Main
  Document are shown in Section \ref{sec:variationaldistributions}. Section
  \ref{sec:computationalcomplexity} shows how we may reduce the computational
  complexity of our variational Bayes procedure. Section 
  \ref{sec:expectedloglikelihood} elaborates on the Variational Bayes expected
  joint log likelihood that is maximised to estimate the penalty multipliers as
  explained in Section 3 of the Main Document. The full procedure of the
  proposed method is shown in Section \ref{sec:fullestimationprocedure}. We
  apply our method to a study on microRNAs in colorectal cancer data 
  \cite[]{neerincx_combination_2018} in Section \ref{sec:appcolon} . Finally, we
  introduce the ridge and lasso versions of the proposed method in Section
  \ref{sec:evbridgelasso}.

@@

  \subsection{Expectation and variance}
  The expectation and variance of $\beta$ are:
  \begin{align*}
  \E[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta \cdot \exp\[
  -\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta=0, \\
  \V[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta^2 \cdot
  \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta -
  \E[\beta]^2\\ &= \frac{\lambda_1^2}{4 \lambda_2^2} + \frac{1}{\lambda_2} -
  \frac{\lambda_1}{2 \lambda_2^{3/2}} \phi \( \frac{\lambda_1}{2 
  \sqrt{\lambda_2}} \) \Phi \(\frac{-\lambda_1}{2 \sqrt{\lambda_2}} \)^{-1}.
  \end{align*}

\subsection{Simulation}
Generating samples from the elastic net prior is not straightforward. \cite{li_bayesian_2010} show that the elastic net prior may be written as a scale mixture of normals, with mixing parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} \tr$: 
\begin{align*}
\beta | \tau & \sim \mathcal{N} \left(0,\frac{1}{\lambda_2} \frac{\tau - 1}{\tau} \right), \\
\tau & \sim \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2}{\lambda_1^2}, \left(1,\infty \right) \right),
\end{align*}
where $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$. So with an efficient truncated gamma sampler available we may sample from the prior using any standard software package that includes a Gaussian sampling routine. More specifically, we require a a sample from the $\mathcal{TG} \left( \frac{1}{2},8 \lambda2/\lambda_1^2, \left(1,\infty \right) \right)$ distribution. The CDF of this distribution is given by:
$$
F_{T}(\tau) =	
\begin{dcases}
0, & \text{if }\tau < 1 \\
1 - \frac{\Phi\(-\frac{\lambda_1}{2} \sqrt{\frac{\tau}{\lambda_2}} \)}{\Phi\(-\frac{\lambda_1}{2 \sqrt{\lambda_2}} \)}, & \text{if } \tau \geq 1.
\end{dcases}
$$

@@

We start with the computation of (\ref{eq:elboomega}). To this end we define:
$$
h(m_i, \omega_i) := \sum_{k=0}^{\infty} (-1)^k \frac{\Gamma(k+m_i}{k + 1}\frac{2k +m_i}{\sqrt{2 \pi \omega_i^3}}\exp\[-\frac{(2k + m_i)^2}{8 \omega_i} \].
$$
Let $\c = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix} \tr$. Furthermore, in the following, all expectations and variances are with respect to $Q^{(t)}$. Next we write out the three terms in (\ref{eq:elboomega}):
\begin{subequations} \label{eq:elboomega2}
	\begin{align}
	\E [\log \L (\y ; \bbeta)] & = \sum_{i=1}^n \log \binom{m_i}{y_i} + \y \tr \X \E [\bbeta] - \m \tr \E \left\{ \log [1 + \exp(\x_i \tr \bbeta)] \right\}, \\
	\E [\log \pi(\bomega | \bbeta)] & = \m \tr \E \left\{ \log [1 + \exp(\x_i \tr \bbeta)] \right\} - n \log 2 - \frac{1}{2} \m \tr \X \E [\bbeta] \\
	& \,\,\,\,\, - \sum_{i=1}^n \log \Gamma (m_i) - \frac{1}{2} \sum_{i=1}^n [ ( \x_i \tr \E [\bbeta] )^2 + \x_i \tr \V [\bbeta] \x_i ] \E [\omega_i] \\
	& \,\,\,\,\,+ \sum_{i=1}^n \E \left\{ \log [ h(m_i, \omega_i) ] \right\}, \\
	\E [\log q(\bomega)] & =  \m \tr \log [1 + \exp(\c)] - n \log 2 - \frac{1}{2} \m \tr \c - \sum_{i=1}^n \log \Gamma (m_i) - \frac{1}{2} \sum_{i=1}^n c_i^2 \E [\omega_i] \\
	& \,\,\,\,\, + \sum_{i=1}^n \E \left\{ \log [ h(m_i, \omega_i) ] \right\},
	\end{align}
\end{subequations}
where all mathematical operations on vectors and matrices are element wise. After combining the terms in (\ref{eq:elboomega2}) and substituting the expectations and variances given in the Main Document, several terms cancel to give:
\begin{subequations} \label{eq:elboomega3}
	\begin{align}
	\E_{Q^{(t)}} & [\log \L (\y ; \bbeta)] + \E_{Q^{(t)}} [\log \pi(\bomega | \bbeta)] - \E_{Q^{(t)}} [\log q(\bomega)] \\ & = \sum_{i=1}^n \log \binom{m_i}{y_i} + ( \y - \frac{1}{2} \m) \tr \X \bmu^{(t)} + \m \tr \left\{ \frac{1}{2} \c^{(t)} - \log [1 + \exp(\c^{(t)})] \right\} \label{eq:elboomega4}\\
	& \,\,\,\,\, + \frac{1}{4} \sum_{i=1}^n \frac{m_i}{c_i^{(t)}} \tanh \(\frac{c_i^{(t)}}{2} \)\left\{ (c_i^{(t)})^2 - ( \x_i \tr \bmu^{(t)} )^2 - \x_i \tr \bSigma^{(t)} \x_i \tr \right\}.
	\end{align}
\end{subequations}

@@

We describe our implementation here for two partitions of the features. To this end, let $(\G_1(1), \dots, \G_1(G_1))$ and $(\G_2(1), \dots, \G_2(G_2))$ denote the two partitions, containing $G_1$ and $G_2$ groups respectively. Furthermore, in the following we assume that the empty sum and empty product evaluate to 0 and 1, respectively. In this two-partition setting we have two penalty multipliers per feature, represented by $\lambda'_{g_1}$ and $\lambda''_{g_2}$, respectively. The generalised frequentist elastic net estimator and corresponding conditional prior are now:
\begin{subequations}\label{eq:multpart}
	\begin{align}
	\hat{\bbeta} &:= \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) -  \frac{\lambda_1}{2} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \sqrt{\lambda'_{g_1} \lambda''_{g_2}} \sum_{\mathclap{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}}} | \beta_j| - \frac{\lambda_2}{2} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \lambda'_{g_1} \lambda''_{g_2} \sum_{\mathclap{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}}} \beta_j^2, \\
	\bbeta | \btau & \sim \prod_{g_1=1}^{G_1} \prod_{g_2=1}^{G_2} \prod_{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}} \mathcal{N} \left(0,\frac{1}{ \lambda'_{g_1} \lambda''_{g_2} \lambda_2} \frac{\tau_j - 1}{\tau_j} \right).
	\end{align}
\end{subequations}
After switching to the parametrisation in \cite{friedman_regularization_2010}, the new penalty multiplier estimates $\blambda'= \begin{bmatrix} \lambda'_1 & \cdots & \lambda'_{G_1} \end{bmatrix} \tr, \blambda''= \begin{bmatrix} \lambda''_1 & \cdots & \lambda''_{G_2} \end{bmatrix} \tr$ are given by:
\begin{subequations}\label{eq:estmultpart}
	\begin{alignat}{2}
	\blambda'^{(k+1)}, \blambda''^{(k+1)} &:= \underset{\blambda', \blambda''}{\argmax} && \Bigg\{ \frac{1}{2} \sum_{g_1=1}^{G_1} |\G_1(g_1)| \log (\lambda'_{g_1}) + \frac{1}{2} \sum_{g_2=1}^{G_2} |\G_2(g_2)| \log (\lambda''_{g_2}) \\ & && - \frac{(1 - \alpha) \lambda}{4} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \lambda'_{g_1} \lambda''_{g_2} d^{(k)}_{g_1 g_2} \Bigg\} \nonumber \\       & \text{subject to } && \prod_{g_1=1}^{G_1} \prod_{g_2=1}^{G_2} (\lambda'_{g_1} \lambda''_{g_2})^{|\G_1(g_1) \cap \G_2(g_2)|} = 1,
	\end{alignat}
\end{subequations}
with the $d_{g_1 g_2}$ terms calculated as $\sum_{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + \alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\)$. The optimisation in (\ref{eq:estmultpart}) is again a convex problem that is easily solved by some numerical optimisation routine. This method naturally generalises to more than two partitions of the data.

@@

  \section{Application to MicroRNAs in colorectal cancer}\label{sec:colorectal}
  In this section we present additional results to Section [! section ref here]
  in the Main Document.

  \subsection{Additional comparisons}

@@

<<>>=
res <- read.table("results/micrornaseq_colorectal_cancer_res3.csv", 
                  stringsAsFactors=FALSE)
medians <- round(apply(res, 2, median), 2)
@
  We compared the estimated multipliers to the estimates by \texttt{GRridge}. 
  Since these results depend on one specific randomisation of the groups, we 
  repeated the procedure 100 times and present the results in Figure 
  \ref{fig:boxplots_micrornaseq_colorectal_cancer_res3_multipliers}. 
  From this figure we see that the estimates for \texttt{gren} are relatively 
  close to one.
  The estimates by \texttt{GRridge} show more variation. We found no strong
  indication of bias, with median estimated penalty parameters 
  \Sexpr{medians[c(1:3)]} for
  \texttt{GRridge} and \Sexpr{medians[c(4:6)]}; \Sexpr{medians[c(7:9)]},
  and \Sexpr{medians[c(10:12)]} for \texttt{gren} with 
  $\alpha \in \{ 0.05, 0.5, 0.95 \}$, respectively.
<<boxplots_micrornaseq_colorectal_cancer_res3_multipliers, fig.cap="Results for the colorectal cancer example with 100 realisations of 3 random groups of features.", out.width="50%", fig.asp=1>>=
@

<<>>=
res <- read.table("results/micrornaseq_colorectal_cancer_res4.csv", 
                  stringsAsFactors=FALSE)
medians <- round(sapply(c("grridge", paste0("gren", 1:3)), function(s) {
  median(as.matrix(res[, grepl(s, colnames(res))]))}), 2)
@
  We repeated the simulation with ten evenly sized groups. Figure 
  \ref{fig:boxplots_micrornaseq_colorectal_cancer_res4_multipliers} presents 
  the estimated penalty multipliers for 100 random 
  splits of the microRNA features into 10 evenly sized groups for 
  \texttt{GRridge} and \texttt{gren} with $\alpha \in \{ 0.05, 0.5, 0.95 \}$. 
  From the Figure we see that \texttt{GRridge} shows more variation around one 
  then the other models. There is no indication of bias: the median penalty 
  multiplier for 
  \texttt{GRridge} is \Sexpr{medians[1]}, while the median multipliers in
  \texttt{gren} are \Sexpr{medians[c(2:4)]} for 
  $\alpha \in \{ 0.05, 0.5, 0.95 \}$, respectively.
<<boxplots_micrornaseq_colorectal_cancer_res4_multipliers, fig.cap="Results for the colorectal cancer example with 100 realisations of 10 random groups of features, plotted (a) with outliers, and (b) without outliers.", out.width="100%", fig.asp=1/2>>=
@

  \subsection{Stability of selection}
  One common criticism on the elastic net, and especially the lasso, is the 
  instability of feature selection. That is, different data sets yield different
  sets of selected features. To investigate the stability of feature selection 
  by \texttt{gren} we created 50 stratified bootstrap samples from the microRNA 
  data \cite[]{neerincx_combination_2018}, where we stratified by the target of 
  prediction, treatment response. 
  For each of these bootstrap samples we estimated the regular elastic
  net and \texttt{gren} for $\alpha \in \{0.05, 0.5, 0.95 \}$, where we selected
  25 features (including the 5 unpenalized covariates). We calculated the size 
  of all ${50}\choose{2}$ intersections of selected features and present the 
  results in Figure [! figure ref here].

@@

  \section{Application to RNAseq in lymph node cancer}
  \texttt{cMCP} is not designed to deal with overlapping groups, so we 
  cross-tabulate the two co-data partitions to create one paritioning with 25 
  groups for \texttt{cMCP}.  
  Additional comparisons to \texttt{SGL}, \texttt{gel} and 
  the overlapping group version of \texttt{gel} \cite[]{zeng_overlapping_2016}
  \cite[]{breheny_group_2015} are found in the SM Section [! section ref here].

>>>>

\documentclass[a4paper,hidelinks]{article}
\usepackage{bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,
rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,
bbm,algorithm,algpseudocode,threeparttable,booktabs,mathtools,
dsfont,parskip}

@@

% functions and operators
\renewcommand{\L}{\mathcal{L}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\argmax}{\text{argmax} \,}
\newcommand{\expit}{\text{expit}}
\newcommand{\erfc}{\text{erfc}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\tr}{^{\text{T}}}
\newcommand{\diag}{\text{diag}}
\newcommand{\KL}{D_{\text{KL}}}
\newcommand{\trace}{\text{tr}}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}

@@

% settings
\graphicspath{{../figs/}}
\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
\setlength{\textwidth}{15cm}
\title{Adaptive group-regularized logistic elastic net regression Supplementary 
material}
\date{\today}
\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: 
\href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Carel F.W. Peeters$^{1}$, 
Aad W. van der Vaart$^{2}$, and \\ Mark A. van de Wiel$^{1,3}$}

@@

  \maketitle

  \noindent
	1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health
	research institute, Amsterdam University medical centers, PO Box 7057, 1007 MB
	Amsterdam, The Netherlands \\*
	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\*
	3. MRC Biostatistics Unit, Cambridge University, Cambridge, United Kingdom

@@

  \section{Introduction}
  This document contains supplementary material to the paper `Adaptive 
  group-regularized logistic elastic net regression'. Section 
  \ref{sec:enetprior} contains extra details on the elastic net prior 
  distribution. Section \ref{sec:objectivefunction} gives the objective 
  function that is maximised in the variational Bayes procedure introduced in
  Section 3 of the Main Document. Section \ref{sec:polyagamma} gives details on
  the P{\'o}lya-Gamma parametrisation introduced in Section 3 of the Main
  Document. The derivations of the variational posterior used in the Main
  Document are shown in Section \ref{sec:variationaldistributions}. Section
  \ref{sec:computationalcomplexity} shows how we may reduce the computational
  complexity of our variational Bayes procedure. Section 
  \ref{sec:expectedloglikelihood} elaborates on the Variational Bayes expected
  joint log likelihood that is maximised to estimate the penalty multipliers as
  explained in Section 3 of the Main Document. The full procedure of the
  proposed method is shown in Section \ref{sec:fullestimationprocedure}. We
  apply our method to a study on microRNAs in colorectal cancer data 
  \cite[]{neerincx_combination_2018} in Section [! section ref here]. Finally, 
  we introduce the ridge and lasso versions of the proposed method in Section
  \ref{sec:evbridgelasso}.

@@

  \subsection{Expectation and variance}
  The expectation and variance of $\beta$ are:
  \begin{align*}
    \E[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta \cdot 
    \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta=0, \\
    \V[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta^2 \cdot
    \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta -
    \E[\beta]^2\\ &= \frac{\lambda_1^2}{4 \lambda_2^2} + \frac{1}{\lambda_2} -
    \frac{\lambda_1}{2 \lambda_2^{3/2}} \phi \( \frac{\lambda_1}{2 
    \sqrt{\lambda_2}} \) \Phi \(\frac{-\lambda_1}{2 \sqrt{\lambda_2}} \)^{-1}.
  \end{align*}

  \subsection{Simulation}
  Generating samples from the elastic net prior is not straightforward. 
  \cite{li_bayesian_2010} show that the elastic net prior may be written as a 
  scale mixture of normals, with mixing parameter 
  $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} \tr$: 
  \begin{align*}
    \beta | \tau & \sim \mathcal{N} \left(0,\frac{1}{\lambda_2} 
    \frac{\tau - 1}{\tau} \right), \\
    \tau & \sim \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2}{\lambda_1^2},
    \left(1,\infty \right) \right),
  \end{align*}
  where $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma 
  distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$. So with
  an efficient truncated gamma sampler available we may sample from the prior 
  using any standard software package that includes a Gaussian sampling routine.
  More specifically, we require a a sample from the $\mathcal{TG} 
  \left( \frac{1}{2},8 \lambda2/\lambda_1^2, \left(1,\infty \right) \right)$ 
  distribution. The CDF of this distribution is given by:
  $$
  F_{T}(\tau) =	
  \begin{dcases}
    0, & \text{if }\tau < 1 \\
    1 - \frac{\Phi\(-\frac{\lambda_1}{2} \sqrt{\frac{\tau}{\lambda_2}} \)}
    {\Phi\(-\frac{\lambda_1}{2 \sqrt{\lambda_2}} \)}, & \text{if } \tau \geq 1.
  \end{dcases}
  $$

@@

  We start with the computation of (\ref{eq:elboomega}). To this end we define:
  $$
  h(m_i, \omega_i) := \sum_{k=0}^{\infty} (-1)^k \frac{\Gamma(k+m_i}{k + 1}
  \frac{2k +m_i}{\sqrt{2 \pi \omega_i^3}}
  \exp\[-\frac{(2k + m_i)^2}{8 \omega_i} \].
  $$
  Let $\c = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix} \tr$. Furthermore, 
  in the following, all expectations and variances are with respect to 
  $Q^{(t)}$. Next we write out the three terms in (\ref{eq:elboomega}):
  \begin{subequations} \label{eq:elboomega2}
    \begin{align}
      \E [\log \L (\y ; \bbeta)] & = \sum_{i=1}^n \log \binom{m_i}{y_i} + 
      \y \tr \X \E [\bbeta] - \m \tr \E \left\{ \log [1 + \exp(\x_i \tr \bbeta)]
      \right\}, \\
      \E [\log \pi(\bomega | \bbeta)] & = \m \tr \E 
      \left\{ \log [1 + \exp(\x_i \tr \bbeta)] \right\} - n \log 2 - 
      \frac{1}{2} \m \tr \X \E [\bbeta] \\
      & \,\,\,\,\, - \sum_{i=1}^n \log \Gamma (m_i) - \frac{1}{2} \sum_{i=1}^n 
      [ ( \x_i \tr \E [\bbeta] )^2 + \x_i \tr \V [\bbeta] \x_i ] 
      \E [\omega_i] \\
      & \,\,\,\,\,+ \sum_{i=1}^n \E 
      \left\{ \log [ h(m_i, \omega_i) ] \right\}, \\
      \E [\log q(\bomega)] & =  \m \tr \log [1 + \exp(\c)] - n \log 2 - 
      \frac{1}{2} \m \tr \c - \sum_{i=1}^n \log \Gamma (m_i) - \frac{1}{2} 
      \sum_{i=1}^n c_i^2 \E [\omega_i] \\
      & \,\,\,\,\, + \sum_{i=1}^n \E \left\{ \log [ h(m_i, \omega_i) ] \right\},
	  \end{align}
  \end{subequations}
  where all mathematical operations on vectors and matrices are element wise. 
  After combining the terms in (\ref{eq:elboomega2}) and substituting the 
  expectations and variances given in the Main Document, several terms cancel to
  give:
  \begin{subequations} \label{eq:elboomega3}
	  \begin{align}
	    \E_{Q^{(t)}} & [\log \L (\y ; \bbeta)] + \E_{Q^{(t)}} 
	    [\log \pi(\bomega | \bbeta)] - \E_{Q^{(t)}} [\log q(\bomega)] \\ 
	    & = \sum_{i=1}^n \log \binom{m_i}{y_i} + ( \y - \frac{1}{2} \m) 
	    \tr \X \bmu^{(t)} + \m \tr \left\{ \frac{1}{2} \c^{(t)} - 
	    \log [1 + \exp(\c^{(t)})] \right\} \label{eq:elboomega4}\\
	    & \,\,\,\,\, + \frac{1}{4} \sum_{i=1}^n \frac{m_i}{c_i^{(t)}} 
	    \tanh \(\frac{c_i^{(t)}}{2} \)\left\{ (c_i^{(t)})^2 - 
	    ( \x_i \tr \bmu^{(t)} )^2 - \x_i \tr \bSigma^{(t)} \x_i \tr \right\}.
    \end{align}
  \end{subequations}

@@

We describe our implementation here for two partitions of the features. To this end, let $(\G_1(1), \dots,\allowbreak \G_1(G_1))$ and $(\G_2(1), \dots, \G_2(G_2))$ denote the two partitions, containing $G_1$ and $G_2$ groups respectively. Furthermore, in the following we assume that the empty sum and empty product evaluate to 0 and 1, respectively. In this two-partition setting we have two penalty multipliers per feature, represented by $\lambda'_{g_1}$ and $\lambda''_{g_2}$, respectively. The generalised frequentist elastic net estimator and corresponding conditional prior are now:
\begin{subequations}\label{eq:multpart}
	\begin{align}
	\hat{\bbeta} &:= \underset{\bbeta}{\argmax} \ell(\y ; \bbeta) -  \frac{\lambda_1}{2} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \sqrt{\lambda'_{g_1} \lambda''_{g_2}} \sum_{\mathclap{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}}} | \beta_j| - \frac{\lambda_2}{2} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \lambda'_{g_1} \lambda''_{g_2} \sum_{\mathclap{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}}} \beta_j^2, \\
	\bbeta | \btau & \sim \prod_{g_1=1}^{G_1} \prod_{g_2=1}^{G_2} \prod_{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}} \mathcal{N} \left(0,\frac{1}{ \lambda'_{g_1} \lambda''_{g_2} \lambda_2} \frac{\tau_j - 1}{\tau_j} \right).
	\end{align}
\end{subequations}
After switching to the parametrisation in \cite{friedman_regularization_2010}, the new penalty multiplier estimates $\blambda'= \begin{bmatrix} \lambda'_1 & \cdots & \lambda'_{G_1} \end{bmatrix} \tr, \blambda''= \begin{bmatrix} \lambda''_1 & \cdots & \lambda''_{G_2} \end{bmatrix} \tr$ are given by:
\begin{subequations}\label{eq:estmultpart}
	\begin{alignat}{2}
	\blambda'^{(k+1)}, \blambda''^{(k+1)} &:= \underset{\blambda', \blambda''}{\argmax} && \Bigg\{ \frac{1}{2} \sum_{g_1=1}^{G_1} |\G_1(g_1)| \log (\lambda'_{g_1}) + \frac{1}{2} \sum_{g_2=1}^{G_2} |\G_2(g_2)| \log (\lambda''_{g_2}) \\ & && - \frac{(1 - \alpha) \lambda}{4} \sum_{g_1=1}^{G_1} \sum_{g_2=1}^{G_2} \lambda'_{g_1} \lambda''_{g_2} d^{(k)}_{g_1 g_2} \Bigg\} \nonumber \\       & \text{subject to } && \prod_{g_1=1}^{G_1} \prod_{g_2=1}^{G_2} (\lambda'_{g_1} \lambda''_{g_2})^{|\G_1(g_1) \cap \G_2(g_2)|} = 1,
	\end{alignat}
\end{subequations}
with the $d_{g_1 g_2}$ terms calculated as $\sum_{\substack{j \in \G_1(g_1) \\ \cap \G_2(g_2)}} \[\bSigma^{(k)}_{jj} + (\bmu^{(k)}_j)^2\]\(1 + \alpha \lambda^{1.5} \sqrt{\frac{1 - \alpha}{8 \chi_j^{(k)}}}\)$. The optimisation in (\ref{eq:estmultpart}) is again a convex problem that is easily solved by some numerical optimisation routine. This method naturally generalises to more than two partitions of the data.

@@

  \section{Application to MicroRNAs in colorectal cancer}\label{sec:colorectal}
  \subsection{Additional comparisons}
  We compared \texttt{gren} to two methods in addition to the ones in the MD: 
  the sparse group lasso (\texttt{SGL}) by \cite{simon_sparse-group_2013} and 
  the group exponential lasso (\texttt{gel}) by \cite{breheny_group_2015}. For
  all methods with a tuning parameter $\alpha$, we set 
  $\alpha \in \{ 0.05, 0.5, 0.95 \}$ and present the AUC on the test set in 
  Figure \ref{fig:lines_micrornaseq_colorectal_cancer_res1_auc}.
<<lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Estimated AUC for the colorectal cancer example.", out.width="100%", fig.asp=1>>=
@

@@

<<>>=
res <- read.table("results/micrornaseq_colorectal_cancer_res2.csv", 
                  stringsAsFactors=FALSE)
medians <- round(apply(res, 2, median), 2)
@
  We compared the estimated multipliers to the estimates by \texttt{GRridge}. 
  Since these results depend on one specific randomisation of the groups, we 
  repeated the procedure 100 times and present the results in Figure 
  \ref{fig:boxplots_micrornaseq_colorectal_cancer_res2_multipliers}. 
  From this figure we see that the estimates for \texttt{gren} are relatively 
  close to one.
  The estimates by \texttt{GRridge} show more variation. We found no strong
  indication of bias, with median estimated penalty parameters 
  \Sexpr{medians[c(1:3)]} for
  \texttt{GRridge} and \Sexpr{medians[c(4:6)]}; \Sexpr{medians[c(7:9)]},
  and \Sexpr{medians[c(10:12)]} for \texttt{gren} with 
  $\alpha \in \{ 0.05, 0.5, 0.95 \}$, respectively.
<<boxplots_micrornaseq_colorectal_cancer_res2_multipliers, fig.cap="Results for the colorectal cancer example with 100 realisations of 3 random groups of features.", out.width="50%", fig.asp=1>>=
@

<<>>=
res <- read.table("results/micrornaseq_colorectal_cancer_res3.csv", 
                  stringsAsFactors=FALSE)
medians <- round(sapply(c("grridge", paste0("gren", 1:3)), function(s) {
  median(as.matrix(res[, grepl(s, colnames(res))]))}), 2)
@
  We repeated the simulation with ten evenly sized groups. Figure 
  \ref{fig:boxplots_micrornaseq_colorectal_cancer_res3_multipliers} presents 
  the estimated penalty multipliers for 100 random 
  splits of the microRNA features into 10 evenly sized groups for 
  \texttt{GRridge} and \texttt{gren} with $\alpha \in \{ 0.05, 0.5, 0.95 \}$. 
  From the Figure we see that \texttt{GRridge} shows more variation around one 
  then the other models. There is no indication of bias: the median penalty 
  multiplier for 
  \texttt{GRridge} is \Sexpr{medians[1]}, while the median multipliers in
  \texttt{gren} are \Sexpr{medians[c(2:4)]} for 
  $\alpha \in \{ 0.05, 0.5, 0.95 \}$, respectively.
<<boxplots_micrornaseq_colorectal_cancer_res3_multipliers, fig.cap="Results for the colorectal cancer example with 100 realisations of 10 random groups of features, plotted (a) with outliers, and (b) without outliers.", out.width="100%", fig.asp=1/2>>=
@

  \subsection{Stability of selection}
  One common criticism on the elastic net, and especially the lasso, is the 
  instability of feature selection. That is, different data sets yield different
  sets of selected features. To investigate the stability of feature selection 
  by \texttt{gren} we created 50 stratified bootstrap samples from the microRNA 
  data \cite[]{neerincx_combination_2018}, where we stratified by the target of 
  prediction, treatment response. 
  For each of these bootstrap samples we estimated the regular elastic
  net and \texttt{gren} for $\alpha \in \{0.05, 0.5, 0.95 \}$, where we selected
  25 features (including the 5 unpenalized covariates). We calculated the size 
  of all $\genfrac(){0pt}{2}{50}{2}$ intersections of selected features and present the 
  results in Figure [! figure ref here].

@@

  \section{Application to RNAseq in lymph node cancer}
  In addition to the results presented in Section [! section ref here] of the 
  MD, we compared \texttt{gren} to the sparse group lasso (\texttt{SGL}) by 
  \cite{simon_sparse-group_2013} and the group exponential lasso (\texttt{gel}) 
  by \cite{breheny_group_2015}. The original group lasso methods are not 
  designed to deal with overlapping groups, so in addition to using the versions
  that allow for overlapping groups, we 
  cross-tabulated the two co-data partitions to create one paritioning with 25 
  groups for \texttt{cMCP} and \texttt{gel}.
  We use $\alpha \in \{0.05, 0.5, 0.95 \}$ for the
  methods that require this extra tuning parameter. The AUCs calculated on the
  validation set are presented in Figure 
  \ref{fig:lines_rnaseq_lymph_node_metastasis_res1_auc}
<<lines_rnaseq_lymph_node_metastasis_res1_auc, fig.cap="Estimated AUC for the lymph node cancer example.", out.width="100%", fig.asp=1>>=
@
